---
title: "Using multiRF for Multi-Omics Integration"
author: "Wei Zhang"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using multiRF for Multi-Omics Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  fig.align = "center",
  message = FALSE,
  warning = FALSE
)
options(width = 90)

library(multiRF)
data("tcga_brca_data", package = "multiRF")
```

## Overview

`multiRF` integrates matched multi-omics data with directed multivariate
random forests. The current workflow has four main stages:

1. fit one forest for each requested response--predictor block direction;
2. construct and fuse sample-level forest weight matrices;
3. separate shared from block-specific signal and cluster both components;
4. optionally calculate inverse minimal depth (IMD), select variables, and
   refit a robust model.

The native C++ engine is the default. It implements multivariate regression,
unsupervised residual forests, all-sample and out-of-bag forest weights,
ordinary and enhanced proximity, weighted candidate-variable sampling, and
depth-based forest IMD. `randomForestSRC` is an optional fallback rather than a
requirement for the standard workflow.

This vignette uses a small deterministic subset of the bundled TCGA BRCA data
so that every evaluated chunk runs quickly. For an analysis, use all eligible
samples, a scientifically justified feature-filtering rule, and more trees
(the workflow default is 500).

## Input data

The input is a **named list** of numeric matrices or data frames. Samples are
rows and features are columns. Every block must have:

- the same number of samples;
- identical, unique sample row names in the same order;
- unique feature names; and
- finite, non-constant numeric features.

Do not pass unordered blocks and rely on their row positions. Align them by
sample identifier before fitting. Missing values and categorical predictors
should be imputed or encoded explicitly before using the native numeric
forest.

```{r prepare-data}
full_dimensions <- data.frame(
  block = names(tcga_brca),
  samples = unname(vapply(tcga_brca, nrow, integer(1))),
  features = unname(vapply(tcga_brca, ncol, integer(1)))
)
knitr::kable(full_dimensions)

demo_n <- 100L
demo_p <- c(gene = 40L, methy = 40L, mirna = 30L)
demo_ids <- rownames(tcga_brca[[1L]])[seq_len(demo_n)]

brca_demo <- lapply(names(tcga_brca), function(block) {
  x <- tcga_brca[[block]]
  x[demo_ids, seq_len(min(demo_p[[block]], ncol(x))), drop = FALSE]
})
names(brca_demo) <- names(tcga_brca)

stopifnot(all(vapply(
  brca_demo,
  function(x) identical(rownames(x), demo_ids),
  logical(1)
)))

knitr::kable(data.frame(
  block = names(brca_demo),
  samples = unname(vapply(brca_demo, nrow, integer(1))),
  features = unname(vapply(brca_demo, ncol, integer(1)))
))
```

`filter_mode = "auto"` is the default and performs an omics-aware dispersion
prefilter before any forest is fitted. Here the bundled matrices are already
curated and then deliberately subset, so the example uses
`filter_mode = "none"`. This input prefilter is distinct from IMD-based
variable selection later in the workflow.

## A complete, fast example

`mrf3()` is the compact user-facing entry point. In this example it fits all
directed forests, constructs the shared and specific components, calculates
raw IMD, and applies the mixture selector without a second forest fit.

```{r fit}
fit <- mrf3(
  dat.list = brca_demo,
  k = NULL,
  ntree = 40,
  filter_mode = "none",
  main_clustering = "similarity",
  run_imd = TRUE,
  run_variable_selection = TRUE,
  variable_selection_args = list(
    method = "mixture",
    signal = "shared",
    level = 0.05,
    re_fit = FALSE
  ),
  return_data = TRUE,
  nthread = 1,
  filter_verbose = FALSE,
  verbose = FALSE,
  seed = 529
)
```

With `k = NULL`, cluster number is selected from the learned representation
without using the clinical annotations introduced later. The mixture rule
keeps this vignette fast. The adaptive OOB filtering rule is described in
[Variable selection](#variable-selection).

```{r summarize-fit}
summary(fit)

fit_overview <- data.frame(
  quantity = c(
    "directed forests", "model top-v", "fused top-v",
    "shared clusters", "selected variables"
  ),
  value = c(
    length(get_models(fit)),
    fit$model_top_v,
    if (is.null(fit$fused_top_v)) "none" else fit$fused_top_v,
    length(unique(get_clusters(fit))),
    sum(vapply(get_selected_vars(fit), length, integer(1)))
  )
)
knitr::kable(fit_overview)
```

The most useful accessors avoid depending on the internal list layout:

```{r accessors}
head(get_clusters(fit))
table(shared_cluster = get_clusters(fit))

specific_cluster_counts <- vapply(
  names(brca_demo),
  function(block) length(unique(get_clusters(fit, block = block))),
  integer(1)
)
specific_cluster_counts

get_top_vars(fit, n = 5)
get_vs_summary(fit)
```

Other commonly used accessors are `get_models()`, `get_connection()`,
`get_weights()`, `get_selected_vars()`, and `get_data()`. Use
`get_clusters(fit, block = "gene")` for a block-specific partition and
`get_clusters(fit, which = "robust")` after running the robust branch.

## Directed forests and connections

A connection is stored as `c(response, predictor)`: the second block supplies
the split variables and the first supplies the multivariate responses. With
three blocks, the default retains all six directed pairwise forests.

```{r connections}
connections <- get_connection(fit)
connection_table <- do.call(rbind, lapply(connections, function(x) {
  data.frame(response = x[[1L]], predictor = x[[2L]])
}))
rownames(connection_table) <- NULL
knitr::kable(connection_table)
```

Set `select_connection = TRUE` only when direction screening is intentionally
part of the analysis. Otherwise all directed models are retained and their
connection scores enter the fusion weights.

For the native multivariate-regression forests, important workflow defaults
are:

| Setting | Default | Meaning |
|:--|:--|:--|
| `ntree` | 500 in `mrf3()` | Trees per directed forest |
| `samptype` | `"swor"` | Approximately 63.2% of samples without replacement |
| `mtry` | `ceiling(p / 3)` | Candidate predictor variables per split |
| `ytry` | `ceiling(q / 3)` | Candidate response variables per split |
| `nsplit` | 10 | Random populated cut points per candidate variable; use 0 to scan all |
| `nodesize` | 5 | Regression terminal-node size setting |
| `forest.wt` | `"all"` | All-sample weights for reconstruction |
| `seed` | 529 | Positive, reproducible seed |

At each node, candidate responses are standardized within that node before
their component split statistics are combined. The response with the largest
component statistic is stored as the multivariate splitting response variable
(MSRV). IMD, however, is a forest-level depth measure: for each tree it uses
`1 / (minimum split depth + 1)` and then averages across trees.

## Forest-weight fusion and shared/specific signal

For each directed model, the workflow starts from its sample-by-sample forest
weight matrix. The default similarity path then:

1. removes self weights;
2. applies model-level top-v truncation and row normalization;
3. normalizes connection scores within each response block;
4. averages the response-level matrices uniformly; and
5. optionally applies fused top-v truncation and row normalization.

This response-stratified fusion is summarized by Eqs. 6--8 in the method
description. A tuned or fixed top-v at least 80% of the sample size is interpreted as no
truncation. `model_top_v = Inf` and `fused_top_v = Inf` also request no
truncation explicitly.

The shared and residual-specific affinities are constructed as
`S = W %*% t(W)`, with the diagonal set to zero. Spectral clustering is the
default similarity backend.

```{r fusion-checks}
W <- fit$reconstruction$W$W_all
S_shared <- fit$shared$clustering$similarity

c(
  minimum_weight_row_sum = min(rowSums(W)),
  maximum_weight_row_sum = max(rowSums(W)),
  maximum_similarity_diagonal = max(abs(diag(S_shared)))
)

fit$shared$frac[, c("data", "shared_frac", "specific_ratio")]
```

For block `k`, the response-specific forest weight matrix gives the shared
reconstruction `Xhat^(k) = W^(k) X^(k)` and the residual
`R^(k) = X^(k) - Xhat^(k)`. An unsupervised forest is then fitted to each
residual matrix to obtain its block-specific weight matrix and similarity.
The reported `shared_frac` is
`1 - ||R^(k)||_F^2 / ||X^(k)||_F^2`. It is a descriptive signal fraction, not
a hypothesis-test p-value.

## Variable selection

Set `run_imd = TRUE` when global raw IMD should be retained without running a
selection branch. `run_variable_selection = TRUE` and
`run_robust_clustering = TRUE` compute the required IMD automatically. The
three IMD selection methods all operate on the unnormalized `[0, 1]` scale:

| Method | Selection rule | When to use it |
|:--|:--|:--|
| `"filter"` | Select above `tau * sd(IMD)`; tune `tau` with true OOB normalized MSE across predictor and response coordinates | Adaptive OOB rule; most computationally intensive |
| `"transformation"` (`"test"`) | Eq. 16 one-sided Student-t transformation with `df = ntree - 1` | Per-tree inferential screening |
| `"mixture"` | Explicit point mass at zero plus two components on positive IMD | Posterior noise-probability selection |

`method = "thres"` is also available when a fixed `se * sd(IMD)` cutoff is
needed. `normalized = FALSE` is the default because selection is performed on
raw IMD. If `re_weights = TRUE`, the retained raw IMD values are used as native
candidate-variable sampling weights during refitting.

The evaluated example used the mixture rule:

```{r selected-variables}
get_vs_summary(fit)
lapply(get_selected_vars(fit), utils::head)
```

IMD values can be inspected independently of the selected set:

```{r plot-imd, fig.width=9, fig.height=6.2}
plot_weights(fit, weight_source = "imd", top = 8)
```

For the adaptive OOB filter, use:

```{r adaptive-filter, eval=FALSE}
vs_filter <- mrf3_vs(
  mod = fit,
  method = "filter",
  signal = "shared",
  k = 3,
  tol = 0.01,
  re_fit = FALSE
)
get_selected_vars(vs_filter)
```

The adaptive OOB filter is defined for shared IMD. With `signal = "all"`, the
workflow uses adaptive filtering for shared IMD and the fixed cutoff for
residual-specific IMD, then returns their union. The transformation and
mixture methods can be applied directly to `signal = "shared"`,
`"specific"`, or `"all"`.

```{r alternative-selectors, eval=FALSE}
vs_transformation <- mrf3_vs(
  fit, method = "transformation", signal = "shared",
  level = 0.05, re_fit = FALSE
)

vs_union <- mrf3_vs(
  fit, method = "mixture", signal = "all",
  level = 0.05, re_fit = FALSE
)
```

Standalone `mrf3_vs()` defaults to `re_fit = TRUE`. The staged workflow does
not refit unless requested, except that `run_robust_clustering = TRUE` must and
therefore does force a final refit.

## Clustering modes and PAM50 annotation

The `main_clustering` argument applies the same representation choice to the
shared and specific branches:

| Mode | Representation | Default clustering backend |
|:--|:--|:--|
| `"similarity"` | `W W^T` from fused or residual weights | Spectral |
| `"proximity"` | Same-terminal-node frequency | PAM |
| `"enhanced_proximity"` | Proximity plus sibling-leaf correction | PAM |

The evaluated fit selected three shared clusters by the eigengap of its learned
similarity. The diagnostic below repeats that calculation over `k = 2, ..., 6`.
It is retrospective to the fit and does not use PAM50 or any other clinical
annotation.

```{r tune-k-diagnostic, fig.width=7.6, fig.height=3.3, dpi=160}
k_diagnostic <- tune_k_clusters(
  S_shared,
  method = "Spectral",
  k_tune = 2:6,
  return_cluster = TRUE,
  plot_k = TRUE
)
stopifnot(k_diagnostic$best_k == length(unique(get_clusters(fit))))
```

### Post hoc comparison with PAM50

PAM50 is used here only after fitting to annotate the unsupervised shared
partition. It was not supplied to feature selection, forest construction,
weight fusion, cluster-number tuning, or clustering. Because PAM50 itself is
derived from gene expression and gene expression is one of the multiRF input
blocks, this is a biological concordance check rather than an independent
ground truth.

```{r pam50-setup, include=FALSE}
clinical_demo <- tcga_brca_clinical[
  match(demo_ids, tcga_brca_clinical$sampleID),
  , drop = FALSE
]
stopifnot(
  !anyDuplicated(tcga_brca_clinical$sampleID),
  identical(as.character(clinical_demo$sampleID), demo_ids)
)

pam50 <- factor(
  clinical_demo$BRCA_Subtype_PAM50,
  levels = c("Basal", "Her2", "LumA", "LumB", "Normal"),
  labels = c(
    "Basal-like", "HER2-enriched", "Luminal A", "Luminal B", "Normal-like"
  )
)
tumor_eval <- clinical_demo$sample_type == "Primary Tumor" & !is.na(pam50)

shared_cluster_raw <- get_clusters(fit)
shared_cluster <- factor(
  shared_cluster_raw,
  levels = sort(unique(shared_cluster_raw)),
  labels = paste0("C", seq_along(sort(unique(shared_cluster_raw))))
)
names(shared_cluster) <- demo_ids

S_pam50 <- S_shared[tumor_eval, tumor_eval, drop = FALSE]
pam50_eval <- droplevels(pam50[tumor_eval])
cluster_eval <- droplevels(shared_cluster[tumor_eval])

pam50_counts <- as.data.frame(
  table(Cluster = cluster_eval, PAM50 = pam50_eval),
  stringsAsFactors = FALSE
)
names(pam50_counts)[3L] <- "samples"
pam50_counts$fraction <- ave(
  pam50_counts$samples,
  pam50_counts$Cluster,
  FUN = function(x) x / sum(x)
)
pam50_counts$label <- ifelse(
  pam50_counts$samples > 0L,
  sprintf("%d\n%.0f%%", pam50_counts$samples, 100 * pam50_counts$fraction),
  ""
)

pam50_k3 <- cluster_metrics(cluster_eval, pam50_eval)
cluster_k5 <- factor(spectral_cl(S_shared, k_tune = 5L)$cl)
pam50_k5 <- cluster_metrics(cluster_k5[tumor_eval], pam50_eval)
pam50_metrics <- rbind(
  `Unsupervised k = 3` = pam50_k3,
  `Fixed k = 5 sensitivity` = pam50_k5
)
pam50_metrics$resolution <- rownames(pam50_metrics)
rownames(pam50_metrics) <- NULL
pam50_metrics <- pam50_metrics[
  , c("resolution", "n", "ari", "jaccard", "nmi", "purity")
]
```

All `r sum(tumor_eval)` primary tumors in the demonstration subset had a PAM50
call. The remaining `r sum(clinical_demo$sample_type == "Solid Tissue Normal")`
adjacent-normal samples had no tumor-subtype label and are excluded from this
comparison. The same native t-SNE coordinates are used in panels a and b;
only the coloring changes. Panel c shows the exact within-cluster composition.

```{r pam50-comparison, fig.width=9.2, fig.height=3.8, dpi=180, fig.cap="Relationship between the multiRF shared partition and PAM50 annotations in the demonstration subset. (a, b) The same native t-SNE embedding of the learned shared similarity for 94 primary tumors, colored by multiRF cluster (a) or PAM50 annotation (b). t-SNE is used only for visualization. (c) Within-cluster PAM50 composition; labels show sample count and row percentage."}
cluster_palette <- c(
  "C1" = "#484878", "C2" = "#7884B4", "C3" = "#B4C0E4"
)
pam50_palette <- c(
  "Basal-like" = "#D24B40",
  "HER2-enriched" = "#CC79A7",
  "Luminal A" = "#3182BD",
  "Luminal B" = "#56B4E9",
  "Normal-like" = "#33B5A5"
)

p_cluster_tsne <- plot_tsne(
  mod = list(dat = S_pam50),
  group = cluster_eval,
  seed = 529,
  max_iter = 250,
  learning_rate = 50,
  label_group = TRUE,
  size = 1.7,
  shape = 21,
  base_size = 8.5,
  main = "multiRF shared clusters"
) +
  ggplot2::scale_fill_manual(values = cluster_palette, drop = FALSE)

p_pam50_tsne <- plot_tsne(
  mod = list(dat = S_pam50),
  group = pam50_eval,
  seed = 529,
  max_iter = 250,
  learning_rate = 50,
  label_group = FALSE,
  position = "bottom",
  size = 1.7,
  shape = 21,
  base_size = 8.5,
  main = "PAM50 annotation"
) +
  ggplot2::scale_fill_manual(values = pam50_palette, drop = FALSE) +
  ggplot2::guides(fill = ggplot2::guide_legend(nrow = 3, byrow = TRUE)) +
  ggplot2::theme(legend.key.height = grid::unit(0.32, "lines"))

p_composition <- ggplot2::ggplot(
  pam50_counts,
  ggplot2::aes(PAM50, Cluster, fill = fraction)
) +
  ggplot2::geom_tile(colour = "white", linewidth = 0.5) +
  ggplot2::geom_text(
    ggplot2::aes(label = label, colour = fraction > 0.45),
    size = 2.5, lineheight = 0.9, show.legend = FALSE
  ) +
  ggplot2::scale_fill_gradientn(
    colours = c("#F7FBFF", "#B4C0E4", "#484878"),
    limits = c(0, 1),
    name = "Within-cluster\nfraction"
  ) +
  ggplot2::scale_colour_manual(values = c("FALSE" = "#272727", "TRUE" = "white")) +
  ggplot2::coord_equal() +
  ggplot2::labs(
    title = "PAM50 composition",
    x = NULL, y = "multiRF cluster"
  ) +
  ggplot2::theme_classic(base_size = 8.5, base_family = "sans") +
  ggplot2::theme(
    axis.line = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_text(angle = 35, hjust = 1),
    plot.title = ggplot2::element_text(face = "bold", size = 9),
    legend.title = ggplot2::element_text(size = 8),
    legend.text = ggplot2::element_text(size = 7.5)
  )

ggpubr::ggarrange(
  p_cluster_tsne,
  p_pam50_tsne,
  p_composition,
  ncol = 3,
  widths = c(1, 1.15, 1.2),
  labels = c("a", "b", "c"),
  font.label = list(size = 10, face = "bold")
)
```

The three-cluster partition captures a coarse organization of PAM50-related
biology rather than a one-to-one recovery of five subtypes: C1 is almost
entirely Basal-like, whereas C2 and C3 divide the luminal-rich samples. The
agreement metrics are reported below. Fixing five clusters on the same learned
similarity is included only as a sensitivity analysis and does not improve
agreement.

```{r pam50-metrics}
knitr::kable(pam50_metrics, digits = 3)
```

The similarity workflow is the default. To compare representations on a full
analysis, keep the data, tree settings, and positive seed fixed:

```{r clustering-modes, eval=FALSE}
fit_similarity <- mrf3(
  dat.list, k = 4, ntree = 500,
  main_clustering = "similarity", seed = 529
)

fit_proximity <- mrf3(
  dat.list, k = 4, ntree = 500,
  main_clustering = "proximity", seed = 529
)

fit_enhanced <- mrf3(
  dat.list, k = 4, ntree = 500,
  main_clustering = "enhanced_proximity", seed = 529
)
```

If `k = NULL`, cluster number is tuned in the downstream clustering stage, as
in the evaluated fit. For a fixed biological hypothesis, supply `k`
explicitly. `tune_k_clusters()` draws its diagnostic on the active graphics
device and restores the prior base-graphics settings afterward.

The ggplot-based helpers can be composed or exported at final figure size. For
example, `ggplot2::ggsave("shared-tsne.pdf", width = 85, height = 75,
units = "mm")` writes a vector figure; use TIFF or PNG at 600 dpi when a raster
file is required. The default `sans` family is portable across R devices.

## Robust and cluster-specific analyses

Robust clustering repeats reconstruction after feature selection. It is an
optional, more expensive branch:

```{r robust, eval=FALSE}
fit_robust <- mrf3(
  dat.list,
  k = 4,
  ntree = 500,
  run_imd = TRUE,
  run_variable_selection = TRUE,
  run_robust_clustering = TRUE,
  variable_selection_args = list(
    method = "filter",
    re_weights = TRUE
  ),
  seed = 529
)

table(get_clusters(fit_robust, which = "robust")$shared)
```

Cluster-specific IMD is separate from global IMD and must be requested
explicitly:

```{r cluster-imd, eval=FALSE}
fit_cluster_imd <- mrf3(
  dat.list,
  k = 4,
  ntree = 500,
  run_imd = TRUE,
  run_cluster_imd = TRUE,
  cluster_imd_args = list(min_cluster_size = 15),
  seed = 529
)

get_weights(fit_cluster_imd$cluster_imd, cluster = 1)
```

## Reproducibility and resource use

- Use a non-negative integer `seed` for a reproducible native fit; 529 is the
  package default. A negative seed deliberately requests system entropy.
  Reproducible seeds give the same result across thread counts.
- Set `options(multiRF.nthread = n)` or pass `nthread = n` to control native
  parallel tree construction.
- Keep `samptype = "swor"` for the default workflow. Sampling with replacement
  is supported when it is scientifically intended.
- Forest weights and proximity matrices are `n` by `n`. For large cohorts,
  account for these matrices when choosing worker counts.
- `compact_output = TRUE` reduces retained duplicate objects, but do not use it
  when models will be needed for later custom reconstruction or reclustering.
- Record `packageVersion("multiRF")`, the random seed, connections, tree
  settings, selected top-v values, and clustering arguments with every result.

For a production fit without automatic input prefiltering:

```{r production-template, eval=FALSE}
options(multiRF.engine = "native", multiRF.nthread = 4L)

fit_final <- mrf3_fit(
  dat.list = dat.list,
  ntree = 500,
  samptype = "swor",
  filter_mode = "none",
  main_clustering = "similarity",
  clustering_args = list(shared_k = 4, specific_k = 2),
  top_v_method = "entropy_elbow",
  select_connection = FALSE,
  run_imd = TRUE,
  run_variable_selection = TRUE,
  variable_selection_args = list(
    method = "filter",
    signal = "shared",
    k = 3,
    re_fit = FALSE
  ),
  return_data = TRUE,
  seed = 529
)
```

## Common problems

**Rows are silently mismatched.** They should not be: current validation stops
on unequal or reordered sample identifiers. Align every block explicitly.

**A constant feature becomes non-finite after scaling.** Remove constant and
non-finite columns before fitting, or use an appropriate filtering mode.

**`mrf3_vs()` cannot find data.** Supply `dat.list` directly or fit with
`return_data = TRUE`.

**A response block is absent from reconstruction.** Per-response fusion
requires at least one fitted response-side connection for every input block;
the workflow now stops rather than substituting a global matrix.

**Block names contain underscores.** Current fits retain explicit response and
predictor metadata, so underscores are supported. Avoid parsing model display
names in downstream code; use `get_connection()`.

## Citation

If you use `multiRF`, please cite the method relevant to your analysis:

> Zhang, W., Wang, L., Franzmann, E. J., and Chen, X. S. (2026).
> Multivariate Random Forests for Cross-Modal Multi-Omics Integration.
> *bioRxiv*. [doi:10.64898/2026.06.17.732933](https://doi.org/10.64898/2026.06.17.732933)

> Zhang, W. et al. (2025). An integrative multi-omics random forest framework
> for robust biomarker discovery. *GigaScience*, 14, giaf148.
> [doi:10.1093/gigascience/giaf148](https://academic.oup.com/gigascience/article/doi/10.1093/gigascience/giaf148/8374728)

## Session information

```{r session-info}
packageVersion("multiRF")
sessionInfo()
```
