---
title: "Getting Started with multiRF"
author: "Wei Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with multiRF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4.5,
  fig.align = "center",
  eval = TRUE,
  message = FALSE,
  warning = FALSE
)

library(multiRF)
data("tcga_brca_data")
```

## Overview

`multiRF` provides a workflow for integrating matched multi-omics datasets with
random forests, MRF-style weighting, clustering, and downstream summaries.

The typical workflow is:

1. Prepare a named list of omics matrices with matched samples.
2. Run `mrf3()` to fit the integration model and produce cluster labels.
3. Inspect weights, selected variables, and cluster assignments.
4. Use plotting helpers for survival and embedding diagnostics.

## Method Maps

### Clustering choices

| Setting | Scope | Typical use |
|---------|-------|-------------|
| `similarity` | Global `main_clustering` | Default fused-similarity workflow; best starting point. |
| `proximity` | Global `main_clustering` | Use when RF proximity is the main object of interest. |
| `enhanced_proximity` | Global `main_clustering` | Aggressive proximity-based branch. |
| `PAM` | Shared or specific branch | Medoid-based clustering. |
| `Spectral` | Shared or specific branch | Similarity geometry-based clustering. |

### Variable-selection choices

| Method | Signal model | Typical use |
|--------|-------------|-------------|
| `filter` | Heuristic threshold | Good baseline for a practical first pass. |
| `test` | Significance-style | Hypothesis-testing flavor for weight selection. |
| `mixture` | Signal/noise mixture | When weights separate into background and enriched components. |

### Plotting choices

| Function | Input focus |
|----------|------------|
| `plot_tsne()` | Similarity or reconstruction |
| `plot_umap()` | Similarity or reconstruction |
| `plot_embed()` | Matrix-based quick embedding |
| `plot_network()` | Similarity graph |
| `plot_weights()` | IMD or cluster-IMD weights |
| `plot_circos()` | Adjacency-like structure |
| `plot_km()` | Cluster labels plus clinical |

## Example Data

```{r example-data}
names(tcga_brca)
dim(tcga_brca$gene)
dim(tcga_brca$methy)
dim(tcga_brca$mirna)
```

## Basic Clustering Workflow

```{r fit-main, cache=TRUE}
fit <- mrf3(
  tcga_brca,
  k = 4,
  ntree = 100,
  filter_mode = "none",
  run_imd = TRUE,
  seed = 529
)
```

```{r inspect-fit}
summary(fit)
table(get_clusters(fit))
names(get_weights(fit))
get_top_vars(fit, n = 10)
```

### Comparing clustering backends

```{r clustering-backends, cache=TRUE}
fit_proximity <- mrf3(
  tcga_brca, k = 4,
  main_clustering = "proximity",
  ntree = 100, seed = 529
)

fit_enhanced <- mrf3(
  tcga_brca, k = 4,
  main_clustering = "enhanced_proximity",
  ntree = 100, seed = 529
)
```

### Compare against PAM50

```{r pam50-compare}
common_ids <- intersect(
  names(get_clusters(fit)),
  tcga_brca_clinical$sampleID
)
pam50 <- tcga_brca_clinical$BRCA_Subtype_PAM50[
  match(common_ids, tcga_brca_clinical$sampleID)
]

cluster_metric_matrix(
  list(
    similarity         = get_clusters(fit)[common_ids],
    proximity          = get_clusters(fit_proximity)[common_ids],
    enhanced_proximity = get_clusters(fit_enhanced)[common_ids],
    PAM50              = pam50
  ),
  metric = "ari"
)
```

## Variable Selection

```{r vs-workflow, cache=TRUE}
fit_vs <- mrf3_fit(
  dat.list = tcga_brca,
  clustering_args = list(shared_k = 4, specific_k = 4),
  ntree = 100,
  run_imd = TRUE,
  run_variable_selection = TRUE,
  variable_selection_args = list(
    method = "mixture",
    level = 0.05,
    normalized = TRUE
  ),
  seed = 529
)
```

```{r vs-inspect}
get_selected_vars(fit_vs)
get_vs_summary(fit_vs)
```

```{r vs-methods, cache=TRUE}
vs_test <- mrf3_vs(
  mod = fit_vs, dat.list = tcga_brca,
  method = "test", re_fit = FALSE, level = 0.05
)
vs_mixture <- mrf3_vs(
  mod = fit_vs, dat.list = tcga_brca,
  method = "mixture", re_fit = FALSE, level = 0.05
)

lapply(
  list(test = vs_test, mixture = vs_mixture),
  get_selected_vars
)
```

## Robust Clustering

```{r robust-workflow, cache=TRUE}
fit_robust <- mrf3_fit(
  dat.list = tcga_brca,
  clustering_args = list(shared_k = 4, specific_k = 4),
  ntree = 100,
  run_imd = TRUE,
  run_variable_selection = TRUE,
  run_cluster_imd = FALSE,
  run_robust_clustering = TRUE,
  variable_selection_args = list(method = "mixture"),
  seed = 529
)
```

```{r robust-inspect}
table(get_clusters(fit_robust, which = "robust"))
summary(fit_robust)
```

### Pairwise IMD

```{r pairwise-imd}
pw <- pairwise_imd(fit_robust)
str(pw, max.level = 1)
```

```{r plot-feature-network, fig.width=7, fig.height=7}
plot_network(pw, label = TRUE, vertex.size = 3, vertex.label.cex = 0.4)
```

## Plotting

```{r plot-tsne, fig.width=6, fig.height=5}
plot_tsne(
  mod = fit,
  group = get_clusters(fit),
  source = "specific_shared",
  main = "t-SNE: shared/specific branch"
)
```

```{r plot-network, fig.width=6, fig.height=6}
plot_network(
  dat = fit,
  group = get_clusters(fit),
  source = "specific_shared",
  cutoff = 0.01
)
```

```{r plot-weights, fig.width=10, fig.height=4}
plot_weights(fit_vs, weight_source = "imd", top = 15)
```

```{r plot-km, fig.width=8, fig.height=6}
clinical_ids <- tcga_brca_clinical$sampleID
clusters <- get_clusters(fit)
if (is.null(names(clusters))) {
  names(clusters) <- rownames(tcga_brca[[1]])
}
km_ids <- intersect(names(clusters), clinical_ids)
clinical_sub <- tcga_brca_clinical[match(km_ids, clinical_ids), , drop = FALSE]
keep <- !is.na(clinical_sub$OS.time) & !is.na(clinical_sub$OS)

plot_km(
  test_var = factor(clusters[km_ids][keep]),
  time_var = "OS.time",
  event_var = "OS",
  pheno_mat = clinical_sub[keep, , drop = FALSE]
)
```

## Session Info

```{r session-info}
sessionInfo()
```
